%-------------------------------------------------------------------------------
% cfg
%-------------------------------------------------------------------------------
%
% \file        cfg.tex
% \library     Documents
% \author      Chris Ahlstrom
% \date        2024-04-16
% \update      2024-08-04
% \version     $Revision$
% \license     $XPC_GPL_LICENSE$
%
%     Provides a description of the entities in the cfg66 library, in the
%     "cfg" namespace.
%
%-------------------------------------------------------------------------------

\section{Cfg Namespace}
\label{sec:cfg_namespace}

   This section provides a useful walkthrough of the \texttt{cfg} namespace of
   the \textsl{cfg66} library.
   Here are the classes (or modules) in this namespace:

   \begin{itemize}
      \item \texttt{cfg::appinfo}
      \item \texttt{cfg::basesettings}
      \item \texttt{cfg::comments}
      \item \texttt{cfg::configfile}
      \item \texttt{cfg::history}
      \item \texttt{cfg::inifile}
      \item \texttt{cfg::inimanager}
      \item \texttt{cfg::inisection}
      \item \texttt{cfg::inisections}
      \item \texttt{cfg::memento}
      \item \texttt{cfg::options}
      \item \texttt{cfg::palette}
      \item \texttt{cfg::recent}
   \end{itemize}

\subsection{cfg::appinfo}
\label{subsec:cfg_namespace_appinfo}

   The \texttt{cfg::appinfo} structure encapsulates basic information about
   an application:

   \begin{itemize}
      \item \textsl{cfg::appkind}.
         Indicates if the application is a headless application (such as a
         daemon), a command-line, \textsl{ncurses}, GUI, or test application.
         In some cases it can be useful to know how the application is
         running.
      \item \textsl{App name}.
         Provides the short name of the application, which can be shown
         in warning messages or be used to name the application as a node, for
         example, in \textsl{JACK}.
      \item \textsl{App version}.
         Provides a version number, or optionally, a variant on the
         \textsl{Name} plus the version number.
      \item \textsl{Main config section name}.
         Holds the name of main configuration section (usually present
         only in the "session" or "rc" file).
         The default value is "[Cfg66]".
      \item \textsl{App home configuration directory}.
         Provides the name of the configuration directory for the application,
         such as \texttt{/home/user/.config/appname}.
         Modified by the \texttt{--home} option.
      \item \textsl{Home configuration file}.
         Provides the name of the main configuration file, such as
         \texttt{appname.rc}.
         Modified by the \texttt{--config} option.
      \item \textsl{Client name}.
         Either the short application name or a variant of it, useful in session
         managers, for example.
         An example under \textsl{NSM} would be \texttt{seq66v2.nUKIE}.
      \item \textsl{App tag}.
         The short application name with the version number tacked on.
      \item \textsl{Arg0}.
         Holds the complete path to the executable as determined at run-time.
      \item \textsl{Package name}.
         Provides the name of the package, which could be the same as the
         application or library name.
      \item \textsl{Session Tag}.
         Provides the session name for the application; it can be the
         application name, a modification of the application name, or
         something completely different, such as a session name given by
         a user.
         Useful in long error/warning/info messages.
      \item \textsl{App icon}.
         Provides the base name of the application icon, if not empty.
      \item \textsl{Version text}.
         Reconstructed version information for the application.
      \item \textsl{API engine}.
         Some applications might use various libraries.
         For example, for a MIDI application it might be one of these
         MIDI libraries:
         \texttt{rtmidi},
         \texttt{rtl66},
         or \texttt{portmidi}.
      \item \textsl{API version}.
         Indicates which version of an API is in force.
         In some cases this can be detected at run-time.
      \item \textsl{GUI version}.
         Indicates the GUI or "curses" version, such as \textsl{Qt 6.1}
         or \textsl{Gtkmm 3.0}.
      \item \textsl{Short client name}.
         Similar to \textsl{client name}, but never has anything appended
         to it.
      \item \textsl{Client name tag}.
         Provides the name to show on the console in error/warning/info messages,
         the short client name surrounded by brackets, such as
         \texttt{[seq66v2]}.
   \end{itemize}

   All of these items can be set at once using the \texttt{appinfo}
   constructor that has a large number of parameters.
   In addition, there are a many "free" functions in the \texttt{cfg}
   namespace for setting and getting these values.
   See \texttt{appinfo.hpp} for a summary.

\subsection{cfg::basesettings}
\label{subsec:cfg_namespace_basesettings}

   THe \texttt{cfg::basesettings} class is a base class.
   It provides common settings useful in any application:

   \begin{itemize}
      \item \textsl{File name}.
         Holds the (optional) name of the file that holds the settings data.
      \item \textsl{Ordinal version}.
         Provides a simple integer indicating the version of the file,
         useful in adapting to changes in settings.
      \item \textsl{Modify/unmodify function and "modified" flag}.
         Indicates if the settings have been modified.
      \item \textsl{Config format}.
         A free-form string indicating the format of the data, such as
         \texttt{INI}, \texttt{XML}, or \texttt{JSON}.
         For "66" libraries and applications, the \texttt{INI}
         format is sufficient.
      \item \textsl{Config type}.
         A short string that indicates something about the
         format or content of the file. For example, in Seq66,
         common values were 'rc' and 'usr', with these
         values also representing the file extension.
      \item \textsl{Ordinal version}.
         A simple integer that is incremented each time a change is
         made in the configuration format or values-present
         that isn't detectable during parsing.
      \item \textsl{Comments}.
         An object holding the main comments that describe something
         about the settings file.
      \item \textsl{Is error}.
         Indicated if there was a error during parsing.
      \item \textsl{Error message}.
         If an error occurred, the error message is stored for display.
   \end{itemize}

\subsection{cfg::comments}
\label{subsec:cfg_namespace_comments}

   \texttt{cfg::comments} holds a string describing
   something general about the configuration,
   meant to be included as the text of a \texttt{[comment]} section.
   Provides some setter and getter functions.
   It is a simple class.

\subsection{cfg::configfile}
\label{subsec:cfg_namespace_configfile}

   The \texttt{cfg::configfile} class is an \texttt{abstract base class}.
   It provides some items common to configuration files, including
   an extensive set of functions to parse sections and configuration
   variables in an INI format in a line-by-line fashion.
   The member functions \texttt{parse()} and \texttt{write()}
   are \textsl{pure virtual}, and must be overridden in a derived class.
   A good example is \texttt{cfg::inifile}.

   These are the main externally-accessible values:

   \begin{itemize}
      \item \textsl{File extension}.
         Common values are 'rc' and 'session'.
      \item \textsl{File name}.
         Provides the file name, normally as a full-path file-specification.
      \item \textsl{File version}.
         Provides the current version of the derived configuration file format.
         Set in the constructor of the configfile-derived object,
         and incremented in that object whenever a new way of reading,
         writing, or formatting the configuration file is created.
   \end{itemize}

   Also too many to list, but it includes functions such as
   \texttt{get\_integer()} and \texttt{write\_integer()}.

   These work by having the whole file read into an
   \texttt{std::ifstream}, and then searching the string over and over
   to read all the variables.
   Sounds inefficient, but in practise it is very fast.

   Finally, there are free functions to delete a configuration file and
   to make a copy of a configuration file.

\subsection{cfg::history}
\label{subsec:cfg_namespace_history}

   One of the things not handled so well in \textsl{Seq66} is the
   undo/redo functionality.
   The \texttt{history} template class implements undo/redo using the 
   \texttt{memento} class described below. It follows
   the \textsl{Design Patterns} book (\cite{dpatterns}).
   Also informative is \cite{deque}.
   Also see the \texttt{cfg::memento} class below.

   The heart of the \texttt{history} template is the history list:

   \begin{verbatim}
      std::deque<memento<TYPE>> m_history_list;
   \end{verbatim}

   Member functions are provided to see if history entries are
   undoable/redoable, undo and redo them, check the maximum size of the list,
   etc.
   The \texttt{history.cpp} module provide a small test of history for
   the \texttt{cfg::options} class.

   \index{history\_test}
   The \texttt{history\_test} application provides a more substantive
   test of history.
   It provides an \texttt{cfg::options::container} with a few options, and
   then applies \texttt{undo} and
   \texttt{redo} operations on them, checking the results.

   \textbf{To do}: test range validation.

\subsection{cfg::inifile}
\label{subsec:cfg_namespace_inifile}

   \texttt{cfg::inifile} is derived from the \texttt{cfg::configfile} class.
   It contains a reference to a class that manages multiple
   INI sections: \texttt{cfg::inisections}.
   It overrides the
   \texttt{cfg::configfile::parse()} and
   \texttt{cfg::configfile::write()} functions.
   It provides two protected functions,
   \texttt{parse\_section()} and
   \texttt{write\_section()}.

   \index{ini\_test}
   The \texttt{ini\_test} application sets up a couple of
   \texttt{cfg::inisection::specification} named structures 
   with a long description and a list of values to be stored.
   The \texttt{cfg::stock\_cfg66\_data()} and
   \texttt{cfg::stock\_comment\_data()} specification are defined as well.
   All section specifications are wrapped in a
   \texttt{cfg::inisections::specification} structure
   that is used to create a
   \texttt{cfg::inisections} object and then use it to
   create an INI file for output, exercising \texttt{inifile::write()}.
   The file generated is \texttt{tests/data/fooout.rc}.

   The same setup is used to exercise \texttt{inifile::parse()}.
   The file read is \texttt{tests/data/fooin.rc}.
   The file generated from \texttt{fooin.rc} is
   \texttt{tests/data/fooinout.rc}.
   The file generated from the test's internal data is
   \texttt{tests/data/fooout.rc}.
   The results can be found in the \texttt{tests/data} directory.

   How does it work?

   \begin{verbatim}
       cfg::inisections sections(exp_file_data, "fooout");
   \end{verbatim}

   The \texttt{exp\_file\_data} specification provides
   the configuration type (file extension without the period)
   \texttt{rc},
   the putative file directory
   \texttt{./tests/data},
   the default base name of the configuration file
   \texttt{ini\_test\_session},
   followed by the description of the INI sections object and
   a list of the sections in the INI sections object.

   Once the \texttt{exp\_file\_data} structure is loaded into the
   \texttt{inisection}, we can hook it to an \texttt{inifile} and
   write the file:

   \begin{verbatim}
       cfg::inifile f_out(sections);
       success = f_out.write();
   \end{verbatim}

   The section data items are written to the assembled full
   file-specification, \texttt{./tests/data/fooout.rc}.

   The next test creates an \texttt{inisections} with the base name
   "fooin". An \texttt{inifile} is created from this section and
   is used to parse \texttt{./tests/data/fooin.rc}.

   \begin{verbatim}
        cfg::inisections sections(exp_file_data, "fooin");
        cfg::inifile f_in(sections);
        success = f_in.parse();
   \end{verbatim}

   Once written, we create another \texttt{inifile}, changing it's base-name
   to "fooinout". Then we write it:

   \begin{verbatim}
         cfg::inifile f_inout(sections, "fooinout");  // changes the name
         success = f_inout.write();
   \end{verbatim}

   The files \texttt{./tests/data/fooin.rc} and 
   \texttt{./tests/data/fooinout.rc} should be almost identical.

\subsection{cfg::inimanager}
\label{subsec:cfg_namespace_inimanager}
         
   \texttt{cfg::inimanager} is meant to manage multiple
   \texttt{cfg::inisections} objects.
   Each \texttt{cfg::inisections} holds multiple
   \texttt{cfg::inisection} objects, and represent the
   setting in an \texttt{cfg::inifile}.
   Each INI file is associated with a unique configuration file type,
   such as "rc" or "session", and a section name,

   \texttt{cfg::inimanager} defines a \texttt{section} types, which is a 
   a map with the key being a configuration type, such as "rc", and
   the value being an \texttt{cfg::inisections} object.

   \texttt{cfg::inimanager} also contains a
   \texttt{cli::multiparser} which can map unique option names to
   the INI file and INI sections to which they belong.
   Not all options from an INI file will be useful from the command
   line, though.

   \index{ini\_set\_test}
   The \texttt{ini\_set\_test} application first defines some
   \textsl{global} options
   peculiar to it: "--list", "--read", "--test", and "--write".
   These are \textsl{global} options because they are not associated with
   an INI file or an INI section.
   The \texttt{cfg::inimanager} is created with this list, and populates
   the options. Then two \texttt{cfg::inisection} objects are
   added, one for \texttt{cfg::small\_data} and
   one for \texttt{cfg::rc\_data}.

   The caller can then exercise the global options, including the internal,
   stock options.  The \texttt{--help} option will show all of the global
   options, plus all of the sections and options defined for
   \texttt{cfg::small\_data} and
   \texttt{cfg::rc\_data}.
   The \texttt{--read} and \texttt{write} options accept a file-name and
   process it.

   Note: if something like the following message appears, fix the issue.
   Otherwise, the option cannot be parsed, even if it appears in the
   help text.

   \begin{verbatim}
      [iniset] Couldn't insert <sets-mode,(rc,[misc])>: Change option to a
      unique name
   \end{verbatim}

   Let us walk through some runs of
   \texttt{ini\_set\_test}.

   \begin{verbatim}
      $ ./build/test/ini_set_test --help
   \end{verbatim}

   This command emits color-coded help text showing the options, a brief
   description of each, and the default value of each option.
   The text is extracted from the setup structures, as provided in
   the file \texttt{./tests/small\_spec.hpp}:

   \begin{verbatim}
      inisection::specification small_misc_data
      inisection::specification small_interaction_data
      inisection::specification small_comments (stock comments)
      inisection::specification small_cfg66_data (stock [Cfg66] section)
      inisections::specification small_data
   \end{verbatim}

   The last specification collects all of the others and adds directory and
   descriptions.  Also included are all the specifications in the
   file \texttt{tests/rc\_spec.hpp}, which we will not show here.


   The first part of the help text shows the "global" options, which consist
   of internal stock options plus some additional test options
   defined by \texttt{cfg::options::container s\_test\_options} in
   \texttt{tests/ini\_set\_test.cpp}:
   "--list", "--read", "--write", and "--tests".
   The global options are not associated with either an INI file
   or an INI section.

   Following the global options are specific options associated with an
   INI file and and INI section.  Here's a partial example:

   \begin{verbatim}
      rc:[misc] A number of miscellaneous options
       --manual-ports           Show real system ports, not 'usr' port names.
                                [false]
       --port-naming=v          Port amount-of-label showing. [short]
   \end{verbatim}

   The "rc" represents the INI file, which can have any name, but with an
   extension of \texttt{.rc}.
   For a given program, there can be only one INI file with that extension.
   The "[misc]" represents an INI section in that INI file.

   The setup process is fairly simple once all of the INI specification
   structures have been coded. First, we add the global test options
   to the stock options. Then we add the \texttt{inisections}
   defined in the test "hpp" files.

   \begin{verbatim}
      cfg::inimanager cfg_set(s_test_options);
      bool success = cfg_set.add_inisections(cfg::small_data);
      if (success)
          success = cfg_set.add_inisections(cfg::rc_data);
   \end{verbatim}

   Next, we need to create a \texttt{cli::multiparser} and parse the
   command-line:

   \begin{verbatim}
      cli::multiparser & clip = cfg_set.multi_parser();
      success = clip.parse(argc, argv);
   \end{verbatim}

   Now, not all of the global options have been implemented, so we describe
   only the ones that work.

   \begin{verbatim}
      $ ./build/test/ini_set_test --list --verbose
   \end{verbatim}

   The "verbose" option does nothing, but the "list" option shows
   the current values of all options, and it stars the ones that
   have been modified from the command line: "verbose" and "list".
   We can add a "log" option to dump the list to a file:

   \begin{verbatim}
      $ ./build/test/ini_set_test --log=t.log --list --verbose
   \end{verbatim}

   The next test option:

   \begin{verbatim}
      $ ./build/test/ini_set_test --test
   \end{verbatim}

   That option is currently a minimal implementation. More to come.

   \begin{verbatim}
      $ ./build/test/ini_set_test --read=tests/data/fooin.rc
   \end{verbatim}

   Note that the "=" can be replaced by a space.

   This option
   depends on being able to find the the desired INI sections object in the
   INI manager and finding that it is active (it has entries).
   This INI sections object can then be used to determine what needs to
   be found in the INI file.

   \begin{verbatim}
      $ ./build/test/ini_set_test --write=t.rc
   \end{verbatim}

   This option writes the current settings to \texttt{t.rcv} and
   \texttt{t.small} in the \texttt{tests/data} directory.

   We can combine a number of options to read an INI file, list the
   data read to a log file, and write it to another file:

   \begin{verbatim}
      $ ./build/test/ini_set_test --list --log test.log
              --read tests/data/ini_set_test.rc --write ini-out.rc
   \end{verbatim}

   The output in \texttt{test.log} and \texttt{ini-out.rc}
   should match up well with \texttt{tests/data/ini\_set\_test.rc}.

   One more use case. Copy \texttt{tests/data/ini\_set\_test.rc}
   to a temporary file, \texttt{i.rc}.
   Then edit all of the comments (indicated by the hash mark, '\#')
   so that they are obviously different.
   Then read that file and write it to another temporary file.

   \begin{verbatim}
      $ ./build/test/ini_set_test --read=i.rc --write=j.rc
   \end{verbatim}

   Comparing \texttt{i.rc} and \texttt{j.rc}, one sees that
   \texttt{j.rc} retains the original comments, which ultimately
   came from the specification defined in the application.
   That is, the comments are not modifiable in the 'rc' file.
   (They aren't modifiable in \textsl{Seq66} either.

   In the future it would be good to read in the comments as well
   and store them in the descriptions.

   We can combine a number of options to read an INI file, list the
   data read to a log file, and write it to another file:

   \begin{verbatim}
      $ ./build/test/ini_set_test --list --log test.log
         --read tests/data/ini_set_test.rc --write ini-out.rc
   \end{verbatim}

   The output in \texttt{test.log} and \texttt{ini-out.rc}
   should match up well with \texttt{tests/data/ini\_set\_test.rc}.

\subsection{cfg::inisection}
\label{subsec:cfg_namespace_inisection}

   \texttt{cfg::inisection}  contains a \texttt{specification} structure that
   provides the name of a section, it's description, and
   a container of \texttt{cfg::options}.
   It is represented in a configuration file by a section or tag name
   enclosed in brackets: \texttt{[output-ports]} as an example.
   Each section can also define a file extension
   (e.g. \texttt{.rc}) to indicate its locus.

   It provides a number of functions to return option setting, help text, and
   more.
   Also provided are free functions to return a stock main configuration
   and a stock comments section.

\subsection{cfg::inisections}
\label{subsec:cfg_namespace_inisections}

   The \texttt{inisections.hpp} file has been split,
   and no long declares four classes.

   \texttt{cfg::inisections} defines four types:

   \begin{itemize}
      \item \texttt{specref}.
          \texttt{std::reference\_wrapper<inisection::specification>}.
      \item \texttt{specrefs}.
          \texttt{std::vector<specref>}.
      \item \texttt{sectionlist}.
         \texttt{std::vector<inisection>}.
      \item \texttt{specification}.
         This structure holds the configuration file's extension,
         directory, base name, description, and a vector of
         \texttt{specrefs}.
   \end{itemize}

   \texttt{cfg::inisections} holds a list of \texttt{inisection} objects.
   It represents all of the files, and all of
   the options that are held by the files.
   The options are in a single list,
   and the INI items look them up by name.

   Functions are provided to pass along to the \texttt{cfg::inisection}
   objects and to look up INI sections and options.
   The various parts of a file-name (path, base-name, and file extension)
   are held, and can be used to get the full file-specification for
   reading and writing an INI file.

   There is a lot to this module.
   For now, see the comment in the \texttt{.hpp} and \texttt{.cpp} files.

\subsection{cfg::memento}
\label{subsec:cfg_namespace_memento}

   The \texttt{cfg::memento} template class is a small object that holds one
   copy of a state object.
   It provides these functions:

   \begin{itemize}
      \item \texttt{memento (const TYPE \& s)}.
      \item \texttt{bool set\_state (const TYPE \& s)}.
      \item \texttt{const TYPE \& get\_state () const}.
   \end{itemize}

   The \texttt{cfg::history} class stores a "history list":
   \texttt{std::deque<memento<TYPE>>}.

\subsection{cfg::options}
\label{subsec:cfg_namespace_options}

   The \texttt{cfg::options} holds a number of items needed for the
   specification, reading, and writing of options.
   The options can be read from configuration file or from the command-line.
   These items are nested in the class:

   \begin{itemize}
      \item \textsl{Static data}.
         Flags and numbers are provided to indicate if the option is enabled
         and how it is to be output in a nice format into an INI file.
      \item \textsl{kind}.
         Indicates if the option is boolean, numerical, a filename, list,
         string and some others. This enumeration makes it easier to
         process the option.
      \item \textsl{spec}.
         This nested structure contains these values. For brevity, the
         \texttt{option\_} portion of the name and the type are not shown:
         \begin{itemize}
            \item \texttt{code}.
               Optional single-character name.
            \item \texttt{kind}.
               Is it boolean, integer, string...?
            \item \texttt{cli\_enabled}.
               Normally true; false disables.
            \item \texttt{default}.
               Either a value or "true"/"false".
            \item \texttt{value}.
               The actual value as parsed.
            \item \texttt{read\_from\_cli}.
               Option already set from CLI.
            \item \texttt{modified}.
               Option changed since read/save.
            \item \texttt{desc}.
               A one-line description of option.
            \item \texttt{built\_in}.
               This option is present in all apps.
         \end{itemize}
      \item \textsl{option}.
         The \texttt{cfg::options::option} is a simple pair of the
         name of the option and the \texttt{spec} that describes it.
      \item \textsl{container}.
         The \texttt{cfg::options::container} is a map (dictionary) of
         option \texttt{spec}s keyed bythe name of the option.
   \end{itemize}

   Also specified are the name of the source file and the name of the
   source section in that file.

   Included are quite a number of functions for looking up option values
   and option characteristics.
   Also include are free functions to make options.

   The \texttt{options.cpp} module not only contains many comments explaining
   the module, but also a statically-initialized list of
   default options that any application can use.
   It is also a good example of how to create a list of options.

\subsection{cfg::palette}
\label{subsec:cfg_namespace_}

   The \texttt{cfg::palette} template class can be used to define
   a palette of color code pair with a platform-specific color class
   such as \texttt{QColor} from \textsl{Qt}.
   This is a feature copped from \textsl{Seq66}.

\subsection{cfg::recent}
\label{subsec:cfg_namespace_recent}

   \texttt{recent} provides for the management of a list of recently-used
   items. It is implemented as a deque of strings.

   Included are functions to get the most-recent file-name, a file-name by
   index (with optional removal of the path, for use in menu displays),
   and functions to manage the list.
   This is a feature adapted and extended from \textsl{Seq66}.

%-------------------------------------------------------------------------------
% vim: ts=3 sw=3 et ft=tex
%-------------------------------------------------------------------------------
